### 2024-02-13 output/paper/paper.qmd
format of paper

---
title: "Replication"
author: Yingzhi Zhang
date: today
date-format: long
abstract: "the abstract of the paper"
format: pdf
header-includes: 
- \usepackage{float}
mainfont: Times New Roman
sansfont: Times New Roman
number-sections: true
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(lubridate)
library(opendatatoronto)
library(knitr)
library(kableExtra)

```

# Introduction

# Data 
## Source and Data Collection

## Data Processing

```{r}
#| echo: false
#| eval: false
#| warning: false
#| message: false
#| label: tbl-01
#| tbl-cap: "First ten rows of the cleaned crisis call data"
#| tbl-cap-location: bottom
#| tbl-pos: H

cleaned_data <- read_csv(paste0(gsub("/paper", "", getwd()), "/data/cleaned_data.csv"))
table01_data <- head(cleaned_data, 10)
table_clean <-
  kable(table01_data, format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down"))

table_clean
```

## Data Visualisation
### Crisis type

```{r}
#| echo: false
#| eval: false
#| warning: false
#| message: false
#| label: fig-type
#| fig-cap: "Trend of crisis calls by type over time"
#| fig-pos: H
#| fig-width: 7
#| fig-height: 3

crisis_type_data <- read_csv(paste0(gsub("/paper", "", getwd()), "/data/crisis_type_data.csv"))
crisis_type_graph <-
ggplot(crisis_type_data, aes(x = Year, y = Number_of_crisis_call, color = Type)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(x = "Year",
       y = "Number of Crisis Calls",
       color = "Crisis Type") +
  theme_minimal() +
  scale_colour_brewer(palette = "Paired")

crisis_type_graph
```

### Time period

```{r}
#| echo: false
#| eval: false
#| warning: false
#| message: false
#| label: fig-time
#| fig-cap: "Distribution of crisis calls over the day by type"
#| fig-pos: H
#| fig-width: 7
#| fig-height: 3

time_data <- read_csv(paste0(gsub("/paper", "", getwd()), "/data/time_data.csv"))
time_graph <-
  ggplot(time_data,
         mapping = aes(x = time_of_day, fill = Type))+
  theme_minimal() +
  geom_bar(position = "dodge2")+
  labs(x = "Time of the day", y = "Number of Crisis Call", fill = "Type") +
  scale_fill_brewer(palette = "Paired")

time_graph
```

### Location

```{r}
#| echo: false
#| eval: false
#| warning: false
#| message: false
#| label: fig-location
#| fig-cap: "Number of crisis calls in different neighbourhoods classified by types"
#| fig-width: 8
#| fig-height: 9
#| fig-pos: H

neighbourhood_data <- read_csv(paste0(gsub("/paper", "", getwd()), "/data/neighbourhood_data.csv"))

neighbourhood_figure <-
  neighbourhood_data %>% 
  ggplot(mapping = aes(x = Neighbourhood_name, y = Number)) +
  geom_boxplot() +
  scale_x_discrete(drop = TRUE) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(x = "Neighbourhood Name", y = "Number of Crisis Calls") +
  facet_grid(rows = vars(Type), cols = vars(Location), scales = "free")

neighbourhood_figure

```

# Discussion

\newpage

# References